Covariate analysis of CMC RNASeq data reprocessed with STAR-feature counts workflow with Ensembl v75 (ACC and DLPFC with CQN and SVA normlaisation)

### 2017-04-10 code/Rmd/CMC_ACC_DLPFC_STAR_ENSEMBL_v75_CQN_SVA.Rmd
Minor modifications

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='CMC')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download reprocessed counts (ACC)
COUNT_ACC_ID = 'syn8414424';
ALL_USED_IDs = COUNT_ACC_ID
COUNT_ACC = downloadFile(COUNT_ACC_ID)

@@

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)

# Add in RIN^2 values
COVARIATES$RIN2 = COVARIATES$RIN^2
ContCovariates = c(ContCovariates, 'RIN2')
```
Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, fig.width=8, fig.height=8}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
draw(COVARIATES.CORRELATION$plot, heatmap_legend_side = 'left')
```

@@

PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], MIN_GENE_CPM=0, MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

@@

### Normalisation (custom design)
Since many covariates are correlated, re-normalising COUNTS with covariates identified here in syn8468116. 
1. "Dx.Tissue" is chosen as the primary variable of interest (i.e., covariate selection is conditioned on diagnosis and tissue)
2. Institution, Gender, IntronicRate, RibozeroBatch, PMI, RIN, TranscriptsDetected, Age\\_of\\_Death, TotalReads, IntergenicRate, PercentAligned, LibraryBatch were chosen as fixed effects and Individual_ID is chosen as random effect
```{r iterative.normalisation, results='asis'}
# Primary variable of interest
postAdjustCovars = c('Dx.Tissue', 'Institution', 'Gender', 'IntronicRate', 'RibozeroBatch', 'PMI', 'RIN', 'TranscriptsDetected',
                     'Age_of_Death', 'TotalReads', 'IntergenicRate', 'PercentAligned', 'LibraryBatch');

@@

# Fit linear model using new weights and new design
VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
ADJUSTED.FIT = lmFit(VOOM.GENE_EXPRESSION)

# Residuals after normalisation
RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(ADJUSTED.FIT, CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset)

# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, 
                                              COVARIATES, 
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)

residualSigCovars[["PC_res"]][[2]]$plotData
```
### SVA
```{r sva}
# Get (null) design matrix
MODEL0 = DM1$design[,-grep('Dx.Tissue',colnames(DM1$design))] # Get null model by removing variable of interest
MODEL0 = MODEL0[,linColumnFinder(MODEL0)$indepCols]
MODEL1 = DM1$design
MODEL1 = MODEL1[,linColumnFinder(MODEL1)$indepCols]

# Source modified functions from sva package
source('../R/num.sv.R')
source('../R/irwsva.build.R')
source('../R/f.pvalue.R')

# Calculate the number of surrogate variables
NUM.SV = num.sv(CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset, 
                MODEL1, method = 'be', seed = 123456, B = 30, tol = 1e-22)

# Estimate surrogate variables
SURR.VAR = irwsva.build(CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset,
                        MODEL1, MODEL0, n.sv = NUM.SV, B = 30, tol = 1e-22)$sv

@@

# Re-estimate voom weights
VOOM.SVA.WEIGHTS = voom(PROCESSED_COUNTS$filteredExprMatrix$counts, 
                        design = cbind(MODEL1, SURR.VAR), plot=F,
                        block = COVARIATES$individualIdentifier, 
                        correlation = correlation$cor)

@@

# Residuals after normalisation
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(VOOM.SVA.FIT,
                                                  CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset)
```

### Residual calculation
Calculate weighted residuals and add back "Dx" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = grep("Dx.Tissue", colnames(DM1$design), value = T)
RESIDUAL.SVA.GENE_EXPRESSION = RESIDUAL.SVA.GENE_EXPRESSION + 
  VOOM.SVA.FIT$coefficients[,varsToAddIn] %*% t(VOOM.SVA.FIT$design[,varsToAddIn])
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(RESIDUAL.SVA.GENE_EXPRESSION))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Clustering residual data
```{r decompse.normalise.data2, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.SVA.GENE_EXPRESSION, scale.=T, center = T)

@@

tree = hclust(as.dist(t(RESIDUAL.SVA.GENE_EXPRESSION)))
cols = WGCNA::labels2colors(COVARIATES.tmp);

@@

### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=10, fig.width=15}
# Fit contrast
contrast = makeContrasts(contrasts=c("Dx.TissueControl_ACC-Dx.TissueOther_ACC",
                                     "Dx.TissueControl_ACC-Dx.TissueSCZ_ACC",
                                     "Dx.TissueOther_ACC-Dx.TissueSCZ_ACC",
                                     "Dx.TissueControl_DLPFC-Dx.TissueOther_DLPFC",
                                     "Dx.TissueControl_DLPFC-Dx.TissueSCZ_DLPFC",
                                     "Dx.TissueOther_DLPFC-Dx.TissueSCZ_DLPFC"),
                         levels = colnames(VOOM.SVA.FIT$coefficients))
FIT.CONTR = contrasts.fit(VOOM.SVA.FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

@@

  normalizationType	= 'CQN'
)

@@

# Store surrogate variables
SURR.VAR %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'CMC_ACC_DLPFC_SurVar.tsv', sep = '\t', row.names=F, quote=F)
SURR_OBJ = File('CMC_ACC_DLPFC_SurVar.tsv', name = 'Surrogate Variables', parentId = CODE$properties$id)
annotations(SURR_OBJ) = annotations(CODE)
annotations(SURR_OBJ)$dataSubType = 'surrogateVariables'
SURR_OBJ = synStore(SURR_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store residual gene expression for eQTL analysis
RESIDUAL.SVA.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_eQTLResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
eEXP_OBJ = File('CMC_ACC_DLPFC_eQTLResidualExpression.tsv', 
                name = 'Normalised, covariates and surrogate variable adjusted residual expression (for eQTL)',
                parentId = CODE$properties$id)
annotations(eEXP_OBJ) = annotations(CODE)
annotations(eEXP_OBJ)$dataSubType = 'residualGeneExpForeQTL'
eEXP_OBJ = synStore(eEXP_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

@@

# Store enrichment results
write.table(enrichment.results, file = 'CMC_ACC_DLPFC_EnrichResults.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('CMC_ACC_DLPFC_EnrichResults.tsv', 
                  name = 'Enrichment Analysis Results (Dx.Tissue)', 
                  parentId = CODE$properties$id)
annotations(ENRICH_OBJ) = annotations(CODE)
annotations(ENRICH_OBJ)$dataSubType = 'enrichResults'
ENRICH_OBJ = synStore(ENRICH_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
stopCluster(cl)
```
|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ---------                      |
|  Covariates                                 |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Offset (for gene length and gc content)    |  `r OFFSET_OBJ$properties$id`    |
|  Design Matrix                              |  `r DM_OBJ$properties$id`        |
|  Residual Expression (for eQTL analysis)    |  `r eEXP_OBJ$properties$id`      |
|  Differential Expression                    |  `r DEXP_OBJ$properties$id`      |
|  Enrichment Analysis                        |  `r ENRICH_OBJ$properties$id`    |

>>>>

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='CMC')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```

### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download reprocessed counts (ACC)
COUNT_ACC_ID = 'syn8414424'
ALL_USED_IDs = COUNT_ACC_ID
COUNT_ACC = downloadFile(COUNT_ACC_ID)

@@

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], function(x){
  x = as.numeric(as.character(gsub('[\\,\\%]','',x)))
})

# Add in RIN^2 values
COVARIATES$RIN2 = COVARIATES$RIN^2
ContCovariates = c(ContCovariates, 'RIN2')
```
Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, fig.width=10, fig.height=8}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
tmp = COVARIATES.CORRELATION$ESTIMATE
tmp[COVARIATES.CORRELATION$PVAL > 0.05] = 0
h = Heatmap(tmp, col = colorRamp2(c(-1,0,1), c('blue','white','red')), name = 'AssocEstimate')
draw(h, heatmap_legend_side = 'left')
```

@@

PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], MIN_GENE_CPM=0, MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
pct.pc = PROCESSED_COUNTS$filteredExprMatrix$genes %>%
  dplyr::rename(ensembl_gene_id = genes) %>%
  left_join(Ensemble2HGNC) %>%
  group_by(gene_biotype) %>%
  summarise(fraction  = n()) %>%
  filter(fraction > 100) %>%
  dplyr::mutate(fraction = fraction/length(PROCESSED_COUNTS$filteredExprMatrix$genes[,1]))
```
`r dim(PROCESSED_COUNTS$filteredExprMatrix$genes)[1]` genes are used for the analysis.

Following is the distribution of biotypes
`r kable(pct.pc)`

### Outlier analysis
Detect outlier samples based on their expression (logCPM) pattern
```{r detect.outliers, fig.height=8, fig.width=8, results='asis', cache = FALSE}
indToRemove = c('MSSM_RNA_ACC_BP_24',

                'MSSM_RNA_ACC_375', 'MSSM_RNA_ACC_170', 'MSSM_RNA_ACC_315', 'MSSM_RNA_ACC_250', 'MSSM_RNA_ACC_21', 'MSSM_RNA_ACC_356', 
                'MSSM_RNA_ACC_318', 'MSSM_RNA_ACC_222', 'MSSM_RNA_ACC_220', 'MSSM_RNA_ACC_379', 'MSSM_RNA_PFC_133',

                'MSSM_RNA_PFC_139', 'MSSM_RNA_PFC_163', 'MSSM_RNA_PFC_260', 'MSSM_RNA_PFC_299', 'MSSM_RNA_PFC_361', 'MSSM_RNA_PFC_89', 
                'MSSM_RNA_PFC_319', 'MSSM_RNA_PFC_348', 'MSSM_RNA_PFC_354', 'MSSM_RNA_PFC_60', 'PENN_RNA_PFC_71', 'PENN_RNA_PFC_56')

# Outliers from Menahce's DLPFC analysis syn3493755
# outliers = c('MSSM_RNA_BP_PFC_1', 'MSSM_RNA_PFC_21', 'MSSM_RNA_BP_PFC_3', 'MSSM_RNA_PFC_103', 'MSSM_RNA_PFC_105', 'MSSM_RNA_PFC_115', 'MSSM_RNA_PFC_169',  'MSSM_RNA_PFC_17',
#   'MSSM_RNA_PFC_18', 'MSSM_RNA_PFC_19', 'MSSM_RNA_PFC_213', 'MSSM_RNA_PFC_214', 'MSSM_RNA_PFC_215', 'MSSM_RNA_PFC_216', 'MSSM_RNA_PFC_217', 'MSSM_RNA_PFC_218', 
#   'MSSM_RNA_PFC_219', 'MSSM_RNA_PFC_222', 'MSSM_RNA_PFC_24', 'MSSM_RNA_PFC_252', 'MSSM_RNA_PFC_259', 'MSSM_RNA_PFC_263', 'MSSM_RNA_PFC_267', 'MSSM_RNA_PFC_275',
#   'MSSM_RNA_PFC_307', 'MSSM_RNA_PFC_322', 'MSSM_RNA_PFC_325', 'MSSM_RNA_PFC_27', 'MSSM_RNA_PFC_332', 'MSSM_RNA_PFC_333', 'MSSM_RNA_PFC_334', 'MSSM_RNA_PFC_335',
#   'MSSM_RNA_PFC_340', 'MSSM_RNA_PFC_35', 'MSSM_RNA_PFC_362', 'MSSM_RNA_PFC_372', 'MSSM_RNA_PFC_373', 'MSSM_RNA_PFC_385', 'MSSM_RNA_PFC_393', 'MSSM_RNA_PFC_395',
#   'MSSM_RNA_PFC_49', 'MSSM_RNA_PFC_54', 'MSSM_RNA_PFC_70', 'MSSM_RNA_PFC_76', 'MSSM_RNA_PFC_77', 'MSSM_RNA_PFC_78', 'MSSM_RNA_PFC_88', 'MSSM_RNA_PFC_93',
#   'FAKE_DLPFC_CMC_MSSM_426', 'PENN_RNA_PFC_29', 'FAKE_DLPFC_CMC_PENN_055', 'PENN_RNA_PFC_75', 'PENN_RNA_PFC_110', 'FAKE_DLPFC_CMC_PENN_106', 'FAKE_DLPFC_CMC_PENN_107',
#   'FAKE_DLPFC_CMC_PENN_108', 'MSSM_RNA_PFC_14', 'MSSM_RNA_BP_PFC_21', 'MSSM_RNA_PFC_293', 'MSSM_RNA_PFC_349', 'MSSM_RNA_PFC_353', 'MSSM_RNA_PFC_363', 'MSSM_RNA_PFC_369',
#   'PENN_RNA_PFC_82', 'PENN_RNA_PFC_89', 'PITT_RNA_PFC_1159', 'PITT_RNA_PFC_1296', 'PITT_RNA_PFC_537',  'PITT_RNA_PFC_567', 'PITT_RNA_PFC_871', 'PITT_RNA_PFC_878',
#   'MSSM_RNA_PFC_133', 'MSSM_RNA_PFC_139', 'MSSM_RNA_PFC_210',  'MSSM_RNA_PFC_260', 'MSSM_RNA_PFC_278', 'MSSM_RNA_PFC_299', 'MSSM_RNA_PFC_348', 'MSSM_RNA_PFC_361',
#   'MSSM_RNA_PFC_60', 'PENN_RNA_PFC_56', 'PITT_RNA_BP_PFC_10003', 'PITT_RNA_BP_PFC_1047', 'PITT_RNA_BP_PFC_1086', 'PITT_RNA_BP_PFC_1092', 'PITT_RNA_BP_PFC_1247',
#   'PITT_RNA_BP_PFC_1282', 'PITT_RNA_BP_PFC_1324', 'PITT_RNA_BP_PFC_1374', 'PITT_RNA_BP_PFC_1391', 'PITT_RNA_BP_PFC_686')

# Find principal components of expression to plot
PC <- prcomp(voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E, scale.=T, center = T)

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_') %>%
  dplyr::mutate(label = SampleID)
plotdata$label[!(plotdata$SampleID %in% indToRemove)] = ''

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Institution, shape=Tissue, size=Age_of_Death))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~.)
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p

# Plot abberent distribution of logcpm counts
tmp1 = voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  tidyr::gather(SampleID, logCPM, -ensembl_gene_id) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID') %>%
              tidyr::separate(Dx.Tissue, c('Dx', 'Tissue'), sep = '_'))

p = ggplot(tmp1 %>%
             dplyr::filter(SampleID %in% indToRemove),
           aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'NONE') + facet_grid(Dx+.~Tissue, scale = 'free')
p

indToRetain = setdiff(colnames(PROCESSED_COUNTS$filteredExprMatrix$counts), indToRemove)
PROCESSED_COUNTS$filteredExprMatrix$counts = PROCESSED_COUNTS$filteredExprMatrix$counts[,indToRetain]
COVARIATES = COVARIATES[indToRetain,]

tmp = tidyr::separate(COVARIATES, Dx.Tissue, c('Dx', 'Tissue'), sep = '_') %>%
  group_by(Dx, Tissue) %>%
  summarise(count = n()) %>%
  spread(Tissue, count)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples from `r length(unique(COVARIATES$Individual_ID))` unique individuals

Based on the expression pattern following samples were tagged as outliers: `r paste(indToRemove, collapse = ', ')`

Distribution of samples are:
`r kable(tmp)`

@@

### Normalisation (custom design)
Since many covariates are correlated, re-normalising COUNTS with a custom design matrix. 
1. "Dx.Tissue" is chosen as the primary variable of interest (i.e., covariate selection is conditioned on diagnosis and tissue)
2.  Model:
    Gx = Dx.Tissue + Institution + Gender + IntronicRate + RibozeroBatch + PMI + RIN + TranscriptsDetected +
    Age\_of\_Death + IntergenicRate + MappedReads + PercentAligned + (1/Individual\_ID)    
```{r custom.normalisation, results='asis'}
# Primary variable of interest
postAdjustCovars = c("Dx.Tissue", "Institution", "Gender", "IntronicRate", "RibozeroBatch", "PMI", "RIN",
                     "TranscriptsDetected", "Age_of_Death", "IntergenicRate", "MappedReads", "PercentAligned");

writeLines(paste('Using following covariates in the model:',
                 paste(postAdjustCovars, collapse=', '),
                 'as fixed effects and Individual_ID is chosen as random effect'))

@@

# Fit linear model using new weights and new design
VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
ADJUSTED.FIT = lmFit(VOOM.GENE_EXPRESSION)
```

### SVA
```{r sva}
# Get (null) design matrix
MODEL0 = DESIGN$design[,-grep('Dx.Tissue',colnames(DESIGN$design))] # Get null model by removing variable of interest
MODEL0 = MODEL0[,linColumnFinder(MODEL0)$indepCols]
MODEL1 = DESIGN$design
MODEL1 = MODEL1[,linColumnFinder(MODEL1)$indepCols]

# Source modified functions from sva package
# source('../R/num.sv.R')
source('../R/irwsva.build.R')
source('../R/f.pvalue.R')

# Get residuals from differential expression model
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(ADJUSTED.FIT, VOOM.GENE_EXPRESSION$E)

# Compute actual variance of all principal components
tmp = svd(RESIDUAL.SVA.GENE_EXPRESSION)
actual.var = tmp$d^2/sum(tmp$d^2)

# Compute permuted variance of all principal components
permuted.var = foreach(i = 1:20, .combine = rbind) %dopar% {
  tmp.residual = t(apply(RESIDUAL.SVA.GENE_EXPRESSION, 1, sample, replace = FALSE))
  tmp = svd(tmp.residual)
  permuted.var = tmp$d^2/sum(tmp$d^2)
}
permuted.var = apply(permuted.var, 2, mean)
NUM.SV = sum(actual.var >= permuted.var[1])

# Plot variance components
var.comp = data.frame(component = 1:length(actual.var), 
                      residual = actual.var, 
                      permuted.residual = permuted.var) %>%
  tidyr::gather(data, value, -component)
p = ggplot(var.comp %>% filter(component <= round(NUM.SV*1.2)), aes(x = component, y = value, color = data)) + geom_point()
p = p + geom_hline(yintercept=permuted.var[1], linetype = 'dashed') + xlab('Singular Dimension Index')
p = p + geom_vline(xintercept=NUM.SV, linetype = 'dashed') + ylab('Fraction of Variance Explained')
p = p + theme(legend.position = c(0.8, 0.8), legend.title = element_blank())
p

# Estimate surrogate variables
SURR.VAR = irwsva.build(CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset,
                        MODEL1, MODEL0, n.sv = NUM.SV, B = 3, tol = 1e-22)$sv

@@

# Re-estimate voom weights
VOOM.SVA.WEIGHTS = voom(PROCESSED_COUNTS$filteredExprMatrix$counts, 
                        design = cbind(MODEL1, SURR.VAR), plot=F,
                        block = COVARIATES$Donor_ID, 
                        correlation = correlation$cor)

@@

# Residuals after normalisation
RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(VOOM.SVA.FIT,
                                              CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset)
```

### Residual calculation
Calculate weighted residuals and add back "Dx" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = grep("Dx.Tissue", colnames(VOOM.SVA.FIT$coefficients), value = T)
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  VOOM.SVA.FIT$coefficients[,varsToAddIn] %*% t(VOOM.SVA.FIT$design[,varsToAddIn])
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(RESIDUAL.GENE_EXPRESSION))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Clustering residual data
```{r decompse.normalise.data2, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION, scale.=T, center = T)

@@

tree = hclust(as.dist(t(RESIDUAL.GENE_EXPRESSION)))
cols = WGCNA::labels2colors(COVARIATES.tmp);

@@

### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=10, fig.width=15}
# Post adjusted design matrix
DESIGN = getDesignMatrix(cbind(COVARIATES[,postAdjustCovars,drop=F], SURR.VAR),
                               Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]

# Estimate voom weights
VOOM.WEIGHTS = voom(PROCESSED_COUNTS$filteredExprMatrix$counts, design=DESIGN$design, plot=F)
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
correlation <- parallelDuplicateCorrelation(VOOM.WEIGHTS, block = COVARIATES$Individual_ID)

# Re-calculate voom weights with correlation of random effects
VOOM.WEIGHTS = voom(PROCESSED_COUNTS$filteredExprMatrix$counts, 
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$Individual_ID, 
                    correlation = correlation$cor)

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
FIT = lmFit(VOOM.WEIGHTS)

# Fit contrast
contrast = makeContrasts(contrasts=c("Dx.TissueControl_ACC-Dx.TissueOther_ACC",
                                     "Dx.TissueControl_ACC-Dx.TissueSCZ_ACC",
                                     "Dx.TissueOther_ACC-Dx.TissueSCZ_ACC",
                                     "Dx.TissueControl_DLPFC-Dx.TissueOther_DLPFC",
                                     "Dx.TissueControl_DLPFC-Dx.TissueSCZ_DLPFC",
                                     "Dx.TissueOther_DLPFC-Dx.TissueSCZ_DLPFC"),
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

@@

  normalizationType	= 'CQN, SVA'
)

@@

# Store residual gene expression for eqtl analysis
RESIDUAL.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
nEXP_OBJ = File('CMC_ACC_DLPFC_netResidualExpression.tsv', 
                name = 'Normalised, covariates removed residual expression (for eQTL analysis)', 
                parentId = CODE$properties$id)
annotations(nEXP_OBJ) = annotations(CODE)
annotations(nEXP_OBJ)$dataSubType = 'residualeQTLAnalysis'
nEXP_OBJ = synStore(nEXP_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

@@

# Store enrichment results
write.table(enrichment.results, file = 'CMC_ACC_DLPFC_EnrichResults.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('CMC_ACC_DLPFC_EnrichResults.tsv', 
                  name = 'Enrichment Analysis Results (Dx.Tissue)', 
                  parentId = CODE$properties$id)
annotations(ENRICH_OBJ) = annotations(CODE)
annotations(ENRICH_OBJ)$dataSubType = 'enrichResults'
ENRICH_OBJ = synStore(ENRICH_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
stopCluster(cl)
```
|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ---------                      |
|  Covariates                                 |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Offset (for gene length and gc content)    |  `r OFFSET_OBJ$properties$id`    |
|  Design Matrix                              |  `r DM_OBJ$properties$id`        |
|  Residual Expression (for eQTL analysis)    |  `r nEXP_OBJ$properties$id`      |
|  Differential Expression                    |  `r DEXP_OBJ$properties$id`      |
|  Enrichment Analysis                        |  `r ENRICH_OBJ$properties$id`    |
